---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r LoadingData, message = FALSE, warning = FALSE}

library(data.table)
activityData <- read.csv("activity.csv", stringsAsFactors = FALSE)
activityData$date <- as.POSIXct(activityData$date)
activityData <- data.table(activityData)

```


## What is mean total number of steps taken per day?

```{r TotalStepsData}
stepsTotalsByDay <- activityData[!is.na(steps), sum(steps), by = date]
```

```{r StepsHistogram}
hist(stepsTotalsByDay$V1, xlab = "Steps per day", breaks = 20, 
     main = "Histogram of Steps Taken per Day")
```

The mean number of steps per day is: **`r format(mean(stepsTotalsByDay$V1), scientific = FALSE)`** 
and the median number of steps per day is: **`r median(stepsTotalsByDay$V1)`**.

## What is the average daily activity pattern?

```{r TotalStepsPerDailyTimeIntervalData}
stepsTotalsByTimeInterval <- activityData[!is.na(steps), mean(steps), by = interval]
plot(stepsTotalsByTimeInterval$interval, stepsTotalsByTimeInterval$V1, 
     main = "Average Steps per Time Interval", xlab = "Interval", ylab = "Number of Steps",
     type = "l")

maximumNumberOfSteps <- max(stepsTotalsByTimeInterval$V1)
intervalOfMaximum <- stepsTotalsByTimeInterval[stepsTotalsByTimeInterval$V1 == maximumNumberOfSteps, ]$interval
```

The maximum average number of steps per interval (**`r maximumNumberOfSteps`**) occurs 
in the interval: **`r intervalOfMaximum`**.


## Imputing missing values

The total number of intervals with NA in the steps column is: **`r nrow(activityData[is.na(activityData$steps)])`**.

In order to handle NA values we will replace all intervals with NA in the steps column with
the mean number of steps for that interval rounded to the nearest integer.

```{r ImputingMissingData}

activityDataWithoutNAs <- activityData
for (i in 1:nrow(activityDataWithoutNAs))
{
    if(is.na(activityDataWithoutNAs[i,]$steps)){
        activityDataWithoutNAs[i,]$steps <- 
        as.integer(round(stepsTotalsByTimeInterval[stepsTotalsByTimeInterval$interval == 
                                                       activityDataWithoutNAs[i,]$interval]$V1))
    }
}

```

```{r TotalStepsDataWithMissingImputed}
stepsTotalsByDayWithMissingImputed <- activityDataWithoutNAs[, sum(steps), by = date]
```

```{r StepsHistogramWithMissingImputed}
hist(stepsTotalsByDayWithMissingImputed$V1, xlab = "Steps per day", breaks = 20, 
     main = "Histogram of Steps Taken per Day")
```

The mean number of steps per day when the missing data is imputed is: **`r format(mean(stepsTotalsByDayWithMissingImputed$V1), scientific = FALSE)`** 
and the median number of steps per day is: **`r median(stepsTotalsByDayWithMissingImputed$V1)`**.

As can be seen by comparing to the figures for mean and median above imputing data from the existing data can affect the analysis of the mean and median.

## Are there differences in activity patterns between weekdays and weekends?

```{r WeekdayDifferences}

activityDataWithWeekdays <- activityDataWithoutNAs
activityDataWithWeekdays$Day <- factor(ifelse(weekdays(activityDataWithWeekdays$date) %in% c("Saturday", "Sunday"),"weekend", "weekday"))

activityMeans <- activityDataWithWeekdays[, mean(steps), by = list(interval, Day)]

library(lattice)
xyplot(V1 ~ interval | Day, data = activityMeans, layout = c(1,2), ylab = "Number of steps", xlab = "Interval", type = "l")



```
