---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r LoadingData, message = FALSE}

library(data.table)
activityData <- read.csv("activity.csv", stringsAsFactors = FALSE)
activityData$date <- as.POSIXct(activityData$date)
activityData <- data.table(activityData)

```


## What is mean total number of steps taken per day?

```{r TotalStepsData}
stepsTotalsByDay <- activityData[!is.na(steps), sum(steps), by = date]
```

```{r StepsHistogram}
hist(stepsTotalsByDay$V1, xlab = "Steps per day", breaks = 20, 
     main = "Histogram of Steps Taken per Day")
```

The mean number of steps per day is: `r format(mean(stepsTotalsByDay$V1), scientific = FALSE)` 
and the median number of steps per day is: `r median(stepsTotalsByDay$V1)`.

## What is the average daily activity pattern?

```{r TotalStepsPerDailyTimeIntervalData}
stepsTotalsByTimeInterval <- activityData[!is.na(steps), mean(steps), by = interval]
plot(stepsTotalsByTimeInterval$interval, stepsTotalsByTimeInterval$V1, 
     main = "Average Steps per Time Interval", xlab = "Interval", ylab = "Number of Steps",
     type = "l")

maximumNumberOfSteps <- max(stepsTotalsByTimeInterval$V1)
intervalOfMaximum <- stepsTotalsByTimeInterval[stepsTotalsByTimeInterval$V1 == maximumNumberOfSteps, ]$interval
```

The maximum average number of steps per interval (`r maximumNumberOfSteps`) occurs 
in the interval: `r intervalOfMaximum`.


## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
